Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-15 paper/main.tex
combine figures

\begin{figure}[H]
    \centering
    \begin{minipage}{0.5\textwidth}
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-1.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-2.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \linebreak
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-3.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-4.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}
    \end{minipage}%
    \begin{minipage}{0.49\textwidth}
        \begin{tikzonimage}[width=\textwidth]
            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}
    \end{minipage}
    \caption{
        The figure shows two masks (A,B) which will be compared against each other.
        As we can see, image (B) contains more cells which has resulted from a division process that
        is not yet occurred in image (A).
        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
        By assigning penalties to all highlighted pixels, we would not take into consideration the
        parental relationship.
        This approach is displayed in (D) where pixels of cells that are related have been colored
        in gray.
    }
    \label{fig:mask-difference-metric}
\end{figure}

We can use this procedure to calculate the cost between successive iteration steps
$t_n$ with $t_{n+1}$.
We expect to obtain an exponential growth in the penalty.
This can be explained by to the fact that the growth of the total area is exponential in time.
Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
exponential is again an exponential.
Figure~\ref{fig:parameter-estimates-single-step} shows the total penalty for this approach.
We can see that by accounting for the parental relationship, we get the overall desired effect.
This method allows us to compare results between states that have a different number of cells
inside as long as their origin stems from the a shared initial state.
However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
process this mask as well as the comparator, it is much more computationally demanding.
In cases where the simpler approach of comparing positions of the agents is reasonable, this method
should be preferred.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/penalty-time-flow.png}%
    \caption{
        Total cost between successive iteration steps $t_n,t_{n+1}$.
        The overall shape is expected to have the slope of an exponential function.
        We can clearly see that division events introduce jumps within this relationship.
        In the case where we account for the parental relationship, these jumps are mitigated.
        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
        parental relationship.
    }
    \label{fig:penalty-calculation-cell-division}
\end{figure}

>>>>

\begin{figure}[H]
    \centering
    \begin{minipage}{0.5\textwidth}
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-1.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-2.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \linebreak
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-3.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-4.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}
    \end{minipage}%
    \begin{minipage}{0.49\textwidth}
        \begin{tikzonimage}[width=\textwidth]
            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}
    \end{minipage}
    \caption{
        The figure shows two masks (A,B) which will be compared against each other.
        As we can see, image (B) contains more cells which has resulted from a division process that
        is not yet occurred in image (A).
        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
        By assigning penalties to all highlighted pixels, we would not take into consideration the
        parental relationship.
        This approach is displayed in (D) where pixels of cells that are related have been colored
        in gray.\\
        Total cost between successive iteration steps $t_n,t_{n+1}$.
        The overall shape is expected to have the slope of an exponential function.
        We can clearly see that division events introduce jumps within this relationship.
        In the case where we account for the parental relationship, these jumps are mitigated.
        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
        parental relationship.
    }
    \label{fig:mask-difference-metric}
\end{figure}

We can use this procedure to calculate the cost between successive iteration steps
$t_n$ with $t_{n+1}$.
We expect to obtain an exponential growth in the penalty.
This can be explained by to the fact that the growth of the total area is exponential in time.
Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
exponential is again an exponential.
Figure~\ref{fig:mask-difference-metric} shows the total penalty for this approach.
We can see that by accounting for the parental relationship, we get the overall desired effect.
This method allows us to compare results between states that have a different number of cells
inside as long as their origin stems from the a shared initial state.
However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
process this mask as well as the comparator, it is much more computationally demanding.
In cases where the simpler approach of comparing positions of the agents is reasonable, this method
should be preferred.
